Download data

### 2023-12-05 scripts/download_dependent_variables.Rmd
all queries ran

```{r load packages}
library(tidyverse)
library(BIEN)
library(ridigbio)
library(rgbif)
source('functions.R')
```

```{r import and clean target species}
targets <- read.csv('../data/raw/2024sos_species.csv', na.strings = "") %>% 
  filter(species != 'tridentata') # purshia, artemisia, and larrea !

@@

## we will not model a few species which are widely abundant, hence do not require modelling, and would utilize enormous computational resources. 
```

@@

gbif_taxon_keys <- target_taxa %>% 
  name_backbone_checklist()  %>% 
  filter(!matchType == "NONE") %>% 
  pull(usageKey)

@@

```{r Perform bien queries, eval = F}

bien_recs <- lapply(target_taxa, BIEN_occurrence_species, collection.info = TRUE,
                    observation.type = TRUE, new.world = TRUE)

@@

ob <- lapply(target_taxa, idig_records)
t <- bind_rows(ob)

@@

file.path(p, list.files(p, pattern = 'gdb$', recursive = T))
# for some strange reason this not exactly working! a weird /gdb is being tagged
# onto the end of it... we can manually specify the data beneath
st_layers(dsn = '../data/2022_AIM_Terrestrial/AIMTerrestrial9-1-22.gdb')

AIM_points <- st_read(dsn = '../data/2021_AIM_Terrestrial/AIMTerrestrial9-1-22.gdb', 
        layer = 'TerrADat', quiet = T) %>%
  filter(State %in% c('NV', 'OR', 'CA', 'ID', 'UT')) %>% 
  select(PrimaryKey, geometry = Shape) %>% 
  st_transform(5070)

AIM_points <- AIM_points[st_covered_by(AIM_points, field_off) %>% lengths > 0,]

spp_richness <- st_read(dsn = '../data/2021_AIM_Terrestrial/AIMTerrestrial9-1-22.gdb', 
        layer = 'tblSpecRichDetail', quiet = T) %>% 
  select(SpeciesList, PrimaryKey) %>% 
  right_join(AIM_points, by = 'PrimaryKey')

>>>>

```{r load packages}
library(tidyverse)
library(BIEN)
library(ridigbio)
library(rgbif)
library(sf)
source('functions.R')
```

```{r import and clean target species}
targets <- read.csv('../data/raw/2024sos_species.csv', na.strings = "") %>% 
  filter(!species %in% c('tridentata', 'millefolium', 'dioica', 'negundo')) # purshia, artemisia, and larrea !

@@

## we will not model a few species which are widely abundant, hence do not require modelling, and would utilize enormous computational resources.

## we will add on USDA codes so we can interface with AIM data.

@@

gbif_taxon_keys <- target_taxa[294:364] %>% 
  name_backbone_checklist()  %>% 
  filter(!matchType == "NONE") %>% 
  pull(usageKey)

@@

BIEN is pretty finicky these days and is prone to crashing. There is a paper by Maitner with Enquist, Wilson, Merow etc. about the need for open source devs to be rewarded academically, I take that has a sign of Brian stepping away from it. We will use a for loop to download from BIEN. 
```{r Perform bien queries, eval = F}

for(i in seq_along(target_taxa)){
  print(target_taxa[i])
  dl <- BIEN_occurrence_species(target_taxa[i], collection.info = TRUE, observation.type = TRUE, new.world = TRUE)
  write.csv(dl, file.path('../data/raw/occurrence/bien', paste0(gsub(' ', '_', target_taxa[i]), '.csv')) )#, row.names = FALSE)
}
```

```{r stitch together BIEN queries}

p <- '../data/raw/occurrence/bien'
files <- file.path(p, list.files(p))

lapply(Filter(function(x) R.utils::countLines(x)==1, files), unlink) # delete empty CSV's in directory
files <- file.path(p, list.files(p))

bien_data <-  do.call(rbind, lapply(files, read.csv)) %>% 
  filter(datasource != 'GBIF' & date_collected > '1950-01-01', ! scrubbed_species_binomial %in% c('Acer negundo', 'Prosopis glandulosa')) %>% 
  drop_na(date_collected, latitude) %>% 
  select(scrubbed_species_binomial:datasource, observation_type)

write.csv(bien_data, '../data/raw/occurrence/BIEN_records.csv', row.names = FALSE)
```

@@

ob <- lapply(target_taxa, idig_records)
idig <- bind_rows(ob) %>% 
  filter(datecollected > '1950-01-01')

write.csv(idig, '../data/raw/occurrence/IDIG.csv')

@@

p1 <- '../data/raw/2021_AIM_Terrestrial/AIMTerrestrial9-1-22.gdb'

# for some strange reason this not exactly working! a weird /gdb is being tagged
# onto the end of it... we can manually specify the data beneath
st_layers(dsn = p1)

AIM_points <- st_read(dsn = p1, 
        layer = 'TerrADat', quiet = T) %>%
#  filter(State %in% c('NV', 'OR', 'CA', 'ID', 'UT')) %>% 
  select(PrimaryKey, geometry = Shape) %>% 
  st_transform(5070)

# AIM_points <- AIM_points[st_covered_by(AIM_points, field_off) %>% lengths > 0,]

spp_richness <- st_read(dsn = p1, 
        layer = 'tblSpecRichDetail', quiet = T) %>% 
  select(SpeciesList, PrimaryKey) %>% 
  right_join(AIM_points, by = 'PrimaryKey')
